A data-driven approach to identifying knowledge gaps in marine invertebrates biodiversity

### 2016-11-22 manuscript.Rmd
begin rewrite intro

With increased anthropogenic pressure on ecosystems, it is crucial to become
efficient at documenting biodiversity before species go extinct. This is
especially important in marine ecosystems where an undefined, but large,
fraction of the biodiversity remains to be discovered especially among
invertebrates. However, taxonomic expertise is dwindling, and there are now only
a handful of natural history museums with active marine invertebrates
collections. Museum collections have been recording our knowledge about species
and their geographical ranges, but until recently, this information was
fragmented across institutions and difficult to synthesize. With collection
digitization and other large-scale projects have aimed at gathering biodiversity
information, public databases can now be used to enable synthesis and highlight
knowledge gaps. Here we use these data to synthetic the state of knowledge for
marine invertebrates, and demonstrate that the stage is set to combine these
data sources to inform where future efforts for documenting biodiversity.

>>>>

With increased anthropogenic pressure on ecosystems, it is crucial to become
efficient at documenting biodiversity before species go extinct. This is
especially important in marine ecosystems where an undefined, but large,
fraction of the biodiversity remains to be discovered especially among
invertebrates. One approach would be to build on existing knowledge to identify
gaps and target understudied areas or taxonomic groups. This is currently
difficult because the knowledge is fragmented and not always
accessible. However, recent several large-scale initiatives have been developed
to aggregate and synthesize various facets of biodiversity knowledge. For
instance, the Biodiversity Heritage Library aims at making the biodiversity
litterature openly available, the World Registry of Marine Species provides a
database of the validity and relationships for most marine species taxonomic
names, and iDigBio digitizes and makes available on the web museum
specimen-associated data. Our knowledge about species and their geographical
range comes from:

1. **Faunistic lists** synthesize and evaluate expert knowledge about a specific
   location. While limited in geographical scope, they tend to be accurate and
   comprehensive.
1. **Museum specimens** are physical proofs of the presence of a species at a
   given point in time and space. The identification can be re-assessed, but
   their contribution to biodiversity knowledge is greatly enhanced if
   associated data and metadata are digitized.
1. **DNA barcodes** useful to delineate and identify species. It requires the
   presence of a museum voucher. Only as good as the identification. Has been
   very useful to identify species complexes, especially among marine
   invertebrates.

To assess the level of digitization for biodiversity, we compared two very
comprehensive faunistic lists for the marine fauna of the US, one from the Puget
Sound (West coast) by Kozloff and the Gulf of Mexico checklist (East
coast). These areas are well studied. Overall, we found that xx% of the species
included in these lists were not found in iDigBio, and xx% in GBIF. Microscopic,
poorly studied metaozans were less represented than larger groups, but for the
latter, proportions on the East and West coast were very similar. Most species
are likely to be undersampled as xx% of the species are represented by less than
5 lots.

However, taxonomic expertise is dwindling, and there are now only
a handful of natural history museums with active marine invertebrates
collections. Museum collections have been recording our knowledge about species
and their geographical ranges, but until recently, this information was
fragmented across institutions and difficult to synthesize. With collection
digitization and other large-scale projects have aimed at gathering biodiversity
information, public databases can now be used to enable synthesis and highlight
knowledge gaps. Here we use these data to synthetic the state of knowledge for
marine invertebrates, and demonstrate that the stage is set to combine these
data sources to inform where future efforts for documenting biodiversity.
